---
title: "Assignment 1"
output: html_document
---
#OVERVIEW

For Assignment 1, I chose to work on a dataset behind the article 'How To Spot A Front-Runner On The ‘Bachelor’ Or ‘Bachelorette’ published on May 2017 by FiveThirtyEight. The article looked at 33 seasons of the show to understand if getting a rose in Week 1 of the show was associated with a higher chance of winning the season. Their findings indicated that there wasn't a strong association between the two. 

For this assignment, I would like to look at a smaller subset of their data (only one season) to try to identify if there are any patterns to elimination and procession for the particular season of the show.

Data for the article was scraped from Bachelor Nation Wiki

Article:https://fivethirtyeight.com/features/the-bachelorette/
Data:https://github.com/fivethirtyeight/data/blob/master/bachelorette/bachelorette.csv
Data Dictionary: https://github.com/fivethirtyeight/data/tree/master/bachelorette


#Code 
```{r} #Installing the Packages Required for Transformation and Analysis
install.packages("RCurl")
install.packages("devtools")
install.packages("plyr")
library(devtools)
library(RCurl)
library (plyr)
```
#Getting the data from the site 
```{r}
x <- getURL("https://raw.githubusercontent.com/fivethirtyeight/data/master/bachelorette/bachelorette.csv")
y <- data.frame(read.csv(text=x))
summary(y)
```
#Creating a subset of the data to show only the contestants from Season 13
```{r} Season13 <- subset(y, SEASON==13)
summary(Season13)
```
#Renaming the columns for readability
```{r}
head(rename(y,c("ELIMINATION.1"="ELIMINATED IN WEEK 1?","ELIMINATION.2"="ELIMINATED IN WEEK 2?","ELIMINATION.3"="ELIMINATED IN WEEK 3?","ELIMINATION.4"="ELIMINATED IN WEEK 4?","ELIMINATION.5"="ELIMINATED IN WEEK 5?","ELIMINATION.6"="ELIMINATED IN WEEK 6?", "ELIMINATION.7"="ELIMINATED IN WEEK 7?", "ELIMINATION.8"="ELIMINTAED IN WEEK 8?", "ELIMINATION.9"="ELIMINATED IN WEEK 9?","ELIMINATION.10"="ELIMINATED IN WEEK 10?")))
```
#FURTHER ANALYSIS
In this assignment, I have created a subset of the main dataset to only show us data for season 13. In addition, I have also renamed the column names for improved readability.
 
The next steps would be to explore and further transform the data (the dataset currently has no numeric data) to perform analyses.


