---
title: "Week 5 Assignment"
author: "Atina Karim"
date: "9/26/2020"
output: html_document
---
# OVERVIEW

For this assignment, I created a wide table, with the arrival data for AM West and Alaska, in Postgresql for further transformation and analysis.


# DATA TRANSFORMATION

## 1. Load Libraries
```{r load libraries}
library(tidyverse)
library(DBI)
```

## 2. Connecting to Postgresql
Results are hidden as it contains db access info.
```{r, include=FALSE}
db <- 'data607'
host_db <- 'localhost'
db_port <- '5432'
db_user <- ''
db_password <- ''
con <- dbConnect (RPostgres::Postgres(),dbname = db, host=host_db, port=db_port, user=db_user, password=db_password)
```
## 3. Listing the table in the db
```{r list tables}
dbListTables(con)
```
## 4. Extracting the data from the db
```{r getting data}
Data <- dbGetQuery(con,"Select * from Arrivals")
```

## 5. Putting the data into a dataframe
```{r create a dataframe}
Df <- data.frame(Data)
```


## 6. Converting data from wide to long
```{r wide to long}
df <- Df %>%
  gather(Destination,flights,
         Los.Angeles,
        Phoenix,
        San.Diego,
        San.Francisco,
        Seattle)
df
```
# ANALYSIS

## On Average, which airline has had more on-time flights?
```{r}
OT <- df %>%
  group_by(Airline) %>%
  filter(Delays == 'On-Time'|Delays == "On time" ) %>%
  summarize(average_ot=mean(flights))
 

```
```{r}
ggplot(OT)+
  geom_col(mapping=aes(x=Airline,y=average_ot,fill=Airline))+
ggtitle("On Time Flights by Airline")
  
```
AM West seems to have more on-time arrivals than Alaska.

## Which airline has more on-time arrivals per city?

```{r}
OTCity <- df %>%
  group_by(Airline,Destination) %>%
  filter(Delays == 'On-Time'|Delays == "On time" ) %>%
  summarize(flights)
```
```{r}
ggplot(OTCity)+
  geom_col(mapping=aes(x=Destination,y=flights,fill=Airline))+
ggtitle("On Time Flights by Airline Per City")
```

Looks like AM West has had more on-time arrivals than Alaska in each city as well,except for in SF and Seattle. 

# **PARADOX**

## Does AM West have more delays as well?



```{r delay}
 df %>%
  group_by(Airline) %>%
  filter(Delays == 'Delayed' ) %>%
  summarize(average_ot=mean(flights))
 
```
It looks like AM West has more delays as well overall. What about per city:

```{r delay city}
DelayCity <- df %>%
  group_by(Airline,Destination) %>%
  filter(Delays == 'Delayed' ) %>%
  summarize(flights)
```

```{r}
arrange(DelayCity,flights)
```
Seems like AM West has more delays per city as well.

## **Explanation**
AM West offers more flights than Alaska, which may explain why it has more of both on-time arrivals and delayed arrivals.
```{r}
df %>%
  group_by(Airline) %>%
  summarize(ot_flights=sum(flights))
```