---
title: "Assignment 3"
author: "Atina Karim"
output:
  html_document:
    df_print: paged
---
#QUESTION 1 
```{r q1-GETTING DATA FROM GITHUB}
library(RCurl)
x <- getURL("https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/majors-list.csv")
y <- data.frame(read.csv(text=x))
summary(y)
```
```{r Data and Statistics}
library(stringr)
y$Major[str_detect(y$Major,pattern="\\DATA|\\STATISTICS")]
```

#QUESTION 2

```{r input data into dataframe}
df <- data.frame(c("bell pepper","bilberry","blackberry","blood orange","blueberry","cantaloupe","chili pepper","cloudberry","elderberry","lime"    ,     "lychee"  ,     "mulberry","olive"    ,    "salal berry"))
```

```{r collapse elements in a vector into a string}
x <- cat(paste(df,collapse=","))
```

#QUESTION 3

1.
Same character appearing consecutively thrice.

```{r}
example <- c("aaa","banana")
str_view(example,"(.)\\1\\1")
```
2.

Repeated pair of the same letters (aa), in between the same letter(s)-like a palindrome
```{r}
example <- c("saas","banana")
str_view(example,"(.)(.)\\2\\1")
```

3.
Repeated pair of letters (i.e. **an** in b**anan**a)
```{r}
example <- c("saas","banana")
str_view(example,"(..)\\1")
```
4. 
Same letter appearing alternatively, thrice (example a in (b)'*a*n*a*n*a*')
```{r}
example <- c("hahaha","banana")
str_view(example,"(.).\\1.\\1")
```
5.
Same letter appearing consecutively thrice, followed by a different letter, then the same initial letters appearing thrice again.
```{r}
example <-c("rrrprrrttcuwjsoooo")
str_view(example,"(.)(.)(.).*\\3\\2\\1")
```
#QUESTION 4

1. Starts and ends with the same character:

```{r}
example <- c("roar","eye","gig","grab")
str_view(example,"(.).*\\1")
```
2.Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)

```{r}
example <- c("church","banana","grab")
str_view(example,"(..).*\\1")
```
3.Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)

```{r}
example <- c("eleven","grab")
str_view(example,"(.).*\\1.*\\1")
```

