---
title: "Week 2 - R AND SQL"
output: html_document
---
**INTRODUCTION**
This assignment seeks to analyze the survey results for Movie Ratings[https://forms.office.com/Pages/DesignPage.aspx#FormId=kWzHivHn_0GonDVTstosFy21RPhPWa5LjqJ9wEetESxUN0dRNTRYTk8yUDE1R1dNUVdJM0RWRU9DWS4u].

The survey asked the respondents to rate six recent movies(Knives Out, Little Women, Parasite, Toy Story 4, Jumanji: The Next Level and Once Upon a Time in Hollywood) on a scale of 1-5 (5=Really Liked it, 1= Did not Like it at all). They also had the option to
select 'Did Not Watch' if they hadn't watched the movie. The survey also required the respondents to indicate their preferred genre of film.

In addition to drawing insights about my friends' movie preferences, I would also like to use the findings from this analysis to identify movies they may like despite not having seen them.


#SQL
I used PostgreSQL to create the database and the table:

##Code for Creating the Database in SQL Shell:##
CREATE DATABASE DATA607;

##Code for Creating the table:##
CREATE TABLE MOVIE_RATINGS(
ID int NOT NULL, Name char (255), KnivesOut_Rating int, Once_Upon_a_Time_in_Hollywood_Rating int, Comments varchar(255), PreferredGenre varchar(255));

*Before loading the results file into the table, I removed the value 'Did not watch' from the ratings column*

##Code for populating the table from the survey:
COPY movie_ratings
FROM 'C:\Users\Atina.Karim\Desktop\DATA607\MovieRatings.csv' 
DELIMITER ',' 
CSV HEADER;

#R
##Connecting to PostGreSQL
```{r}
library(DBI)

db <- 'data607'  #provide the name of your db

host_db <-  'localhost'

db_port <- '5432'  

db_user <- 'postgres'

db_password <- ''

con <- dbConnect(RPostgres::Postgres(), dbname = db, host=host_db, port=db_port, user=db_user, password=db_password)  
```
##Listing the tables in the database
```{r list tables}
dbListTables(con) 
```

```{r}
movie <- dbGetQuery(con,"Select * from move_ratings")
```
## Looking at the summary statistics for our table
```{r list tables}
summary(movie)
```

#WHICH MOVIE RECEIVED THE HIGHEST AVERAGE RATING?
By looking at the averages and medians, we can see that Parasite was rated the highest, followed by Little Women, Knives Out, Toy Story 4, Once Upon a Time in Hollywood and then Jumanji.

#WHICH MOVIE DID ALL THE RESPONDENTS WATCH?
Knives Out is the only movie that all the respondents seem to have seen (no NAs).


##What is the preferred genre of film among my friends?
```{r}
library(ggplot2)
barplot(table(movie$preferredgenre),
main='Genre',
col='red3')
```

Thriller seems to be the preferred genre of film among my friends. This may also provide insight into why Parasite (which is a thriller) was so highly rated among the respondents. This may also explain why Knives Out (also a thriller) was watched by all (and also highly rated).

However, I know more than 2 respondents put down thriller as their preferred genre. So why is the chart only showing a frequency of 2? This maybe because of two of the respondents put down other variations of thriller (such as 'suspense')as their preferred genre so for the purpose of this analysis, I will change the genre for this respondent to Thriller.

```{r}
movie$preferredgenre  <- ifelse(movie$preferredgenre   == "Old Hollywood, 70-80s romantic comedies, Thai horror, Cantonese thrillers"
, "Thriller", movie$preferredgenre)
```
```{r}
movie$preferredgenre  <- ifelse(movie$preferredgenre   == "Suspense","Thriller", movie$preferredgenre)
```
Lets see if this changes the numbers in the barplot.
```{r}
library(ggplot2)
barplot(table(movie$preferredgenre),
main='Genre',
col='red3')
```
The number for thriller has now jumped up to 4.

#Recommendation Analysis
Among my friends who liked Thriller, I would like to see how many of them watched Little Women (movie with the most missing ratings)

```{r}
library(sqldf)
LittleWomen <- sqldf("SELECT count(LittleWoman_Rating)
                FROM movie
               WHERE preferredgenre='Thriller'")
glimpse(LittleWomen)
```

Looks like only one of them did not watch Little Women. Now I would like to see which friend this is:

```{r}
DidNotWatchLittleWomen <- sqldf("SELECT NAME
                FROM movie
               WHERE preferredgenre='Thriller'
               AND LittleWoman_Rating IS NULL")
```

Looks like Faizaa Fatima did not watch Little Women.Before I recommend the movie to her, I would like to see what other thriller lovers rated the movie on an average:

```{r}
AverageLW<- sqldf("Select avg(LittleWoman_Rating)
                  FROM movie 
                  WHERE    preferredgenre='Thriller'")
glimpse(AverageLW)
```
Looks like it got a rating of 5 on an average from those who like Thriller. So I would recommend Little Women to Faizaa as well.

