---
title: "Week 9 assignment"
author: "Atina Karim"
date: "10/23/2020"
output: 
 html_document:
  toc: true
  toc_float: true
  toc_depth: 3
  df_print: paged
  theme: united 
---
# *New York Time Movie Reviews*

I am interested in seeing the NYT movie reviews for movies that were released in 2019. I am particularly interested in seeing how many of the movies that were reviewed, were 'critic picks'.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Loading librarie
```{r , message=FALSE}
library(jsonlite)
library(dplyr)
library(stringr)
library (ggplot2)
```
## Fetching the data from the Movie API
I identified the Movie API from the NYT developer site and got my API key for it. Since I am only interested in looking at the reviews for movies that were released in 2019, I modified the API to only include movies that had an opening date between 2019-01-01 to 2020-01-01.
```{r}

url <- "https://api.nytimes.com/svc/movies/v2/reviews/search.json?opening-date=2019-01-01;2020-01-01"
key <- "OkVf8SLjqbsAbQAvVbiJBn6yRY7azROI"
addurl <- paste0(url, "&api-key=")
# fetched using json + key call
data <- fromJSON(paste0(addurl, key))
summary(data)


```
## Seeing the results as a dataframe
```{r, warning=FALSE}
df <- data$results
knitr:: kable (df)
```
I would like to clean the results up a bit and get rid of the last two columns,
```{r}
df = subset(df,select=-c(link,multimedia))
```
## How many of the movies that released in 2019, were reviewed?
```{r}
nrow(df)
```
## Out of the 20 movies reviewed, how many of them were picked by critics
The critics pick column indicates if a movie was picked by critics or not (1=yes,0=no).
```{r}
critics_pick <- df %>%
  filter(critics_pick==1)
nrow(critics_pick)
```
 Out of the 20 movies released in 2019 that were reviewed, only 3 were critic picks.
 
## What were the movies that were picked by the critics?
 
```{r}
knitr::kable(critics_pick$display_title)
```



